---
title: "Diversity in Pmt2 and S1"
author: "Haydeé"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

We analysed the two metagenomes Pmt2 and S1 using phyloseq. First we process the biom file and subset by Bacteria. 

```{r, include=FALSE}
library("phyloseq")
library("ggplot2")
library("RColorBrewer")
library("patchwork")
library("tidyverse")
```


```{r, include=FALSE}
raw_metagenomes <- import_biom("~/R_sites/metagenomes_/biom-files/metagenomes_Pmt_S1.biom")
class(raw_metagenomes)


raw_metagenomes@tax_table@.Data <- substring(raw_metagenomes@tax_table@.Data, 4)


colnames(raw_metagenomes@tax_table@.Data)<- c("Kingdom", "Phylum", "Class", "Order", "Family", "Genus", "Species")


unique(raw_metagenomes@tax_table@.Data[,"Kingdom"])

raw_metagenomes <- subset_taxa(raw_metagenomes, Kingdom == "Bacteria")
```

```{r, include=FALSE}
phylums <- sort(unique(raw_metagenomes@tax_table@.Data[,"Phylum"]))
Genus <- sort(unique(raw_metagenomes@tax_table@.Data[,"Genus"]))
```

There are `r length(phylums) -1` unique Phylums and `r length(Genus)-1` unique Genus across the two metagenomes.




The most abundant Phylums are in the following table.

```{r}
summary(raw_metagenomes@tax_table@.Data== "")
raw_metagenomes <- subset_taxa(raw_metagenomes, Genus != "")
relative_metagenomes <- transform_sample_counts(raw_metagenomes, function(x) x*100 / sum(x) )
relative_phylums <- tax_glom(relative_metagenomes, taxrank = 'Phylum')
relative_phylums_df <- psmelt(relative_phylums)
relative_phylums_df[order(-relative_phylums_df$Abundance),][1:11,]
```
```{r}
relative_phylums_df2 <- relative_phylums_df[-c(1,4,5)]
rel_phyls <- spread(relative_phylums_df2, Sample, Abundance)
```

```{r}
rel_phyls
```


```{r}
rel_phyls[order(-rel_phyls$Pmt2_S287), ]
```

```{r}
rel_phyls[order(-rel_phyls$S1_S286), ]
```


```{r}
rel_phyls %>% filter(rel_phyls$Pmt2_S287 > 2.2*rel_phyls$S1_S286)
```

```{r}
rel_phyls %>% filter(1.5*rel_phyls$Pmt2_S287 < rel_phyls$S1_S286)
```

```{r}
rel_phyls %>% filter(abs(rel_phyls$Pmt2_S287 - rel_phyls$S1_S286) < 0.001)
```


The deepth of the samples are the following:

```{r}
sample_sums(raw_metagenomes)
```





```{r}
relative_genus <- tax_glom(relative_metagenomes, taxrank = 'Genus')
relative_genus_df <- psmelt(relative_genus)
relative_genus_df$Genus[relative_genus_df$Abundance < 1] <- "Genus < 1% abundance"
```

Genus with abundance > 1% in Pmt2.

```{r}
relative_genus_df %>% filter(Sample=="Pmt2_S287" & Abundance >= 1)%>% count(Genus)
```

Genus with abundance > 1% in Topaz.

```{r}
relative_genus_df %>% filter(Sample=="S1_S286" & Abundance >= 1)%>% count(Genus)
```

```{r}
relative_genus_df %>% filter(Sample=="Pmt2_S287" & Abundance < 1)%>% count(Genus)
```


```{r}
relative_genus_df %>% filter(Sample=="S1_S286" & Abundance < 1)%>% count(Genus)
```

If we subset just by Actinobacteria, we obtain the following.

```{r, include=FALSE}
actinos <- subset_taxa(raw_metagenomes, Phylum == "Actinobacteria")
relative_actinos <- transform_sample_counts(actinos, function(x) x*100 / sum(x) )

relative_actinos_genus <- tax_glom(relative_actinos, taxrank = 'Genus')
relative_actinos_df <- psmelt(relative_actinos_genus)
relative_actinos_df$Genus[relative_actinos_df$Abundance < 1] <- "Genus < 1% abundance"


```

```{r}
relative_actinos_df2 <- psmelt(relative_actinos_genus)
length(sort(unique(relative_actinos_df2[,"Genus"])))
```

There are `r length(sort(unique(relative_actinos_df2[,"Genus"]))) ` unique genus across Actinobacteria.



There are 17 unique Genus in sample Pmt2 with abundance >= 1.

```{r}
relative_actinos_df %>% filter(Sample=="Pmt2_S287" & Abundance >= 1)%>% count(Genus)
```

There are 19 unique Genus in sample S1 with abundance >= 1.

```{r}
relative_actinos_df %>% filter(Sample=="S1_S286" & Abundance >= 1)%>% count(Genus)
```

And in total, there are 22 Genus with abundance over than 1%.

```{r}
relative_actinos_df %>% filter(Abundance >= 1)%>% count(Genus)
```

There are 205 Genus with abundance less than 1% in sample Pmt2 and 203 in Sample S1.
```{r}
relative_actinos_df %>% group_by(Sample) %>%filter(Abundance < 1)%>% count(Genus)
```

```{r}
relative_genus_df %>% filter(Sample=="Pmt2_S287" & Phylum=="Actinobacteria" & Abundance >= 1)%>% count(Genus)
```


```{r}
relative_genus_df %>% filter(Sample=="S1_S286" & Phylum=="Actinobacteria" & Abundance >= 1)%>% count(Genus)
```


```{r}
relative_genus_df %>% filter(Sample=="Pmt2_S287" & Phylum=="Actinobacteria" & Abundance < 1)%>% count(Genus)
```


```{r}
relative_genus_df %>% filter(Sample=="S1_S286" & Phylum=="Actinobacteria" & Abundance < 1)%>% count(Genus)
```

